\newpage
\chapter*{\centering Abstract}
{

 In this project, we investigate the performance of various deep learning models for handwritten English character recognition. Specifically, we explore convolutional neural network (CNN), CNN with gated recurrent units (GRU), CNN with long short-term memory (LSTM), and LSTM models for this task. We also develop and utilize a dataset preparation pipeline to preprocess and augment the available data to improve model performance. Through comprehensive experimentation and performance analysis, we demonstrate the effectiveness of these models for handwritten English character recognition, with the CNN+LSTM model achieving the highest accuracy of all the models tested. Our work contributes to the field of handwriting recognition and provides valuable insights into the strengths and weaknesses of different deep learning models for this task.

The task of handwritten English character recognition has various practical applications, including but not limited to OCR (optical character recognition), document digitization, and text recognition for mobile devices. Our work on developing and analyzing deep learning models for this task is highly relevant and useful in these domains.

Furthermore, our dataset preparation pipeline can be used as a blueprint for preparing and augmenting datasets for other tasks in computer vision and deep learning. This can save significant time and effort in dataset preparation, especially for tasks involving limited amounts of data.



}

 